---
output:
  html_document: default
  pdf_document: default
---
# Regression and model validation

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods.
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using.

This week we understood data wrangling, perform explanatory examination and fit a simple linear model to the data.

## Reading data

Let's read the data
```{r readdata,echo=TRUE,results='hide',message=FALSE,warning=FALSE}
library(dplyr)
learning2014 <-readxl::read_excel("~/IODS-project/data 2/learning2014.xlsx") %>%

  mutate_at(vars(gender), factor)
```

```{r datastructure}
str(learning2014)
```
# the data has 166 observations and 7 variables
# Mostly likert scale (1-5) variables

# Includes also Age (positive integers) and gender (as a two level factor: F and M)

# Attitude = Global attitude toward statistics ~Da+Db+Dc+Dd+De+Df+Dg+Dh+Di+Dj

# Points=Exam points (0-33)

```{r datastructure}
dim(learning2014)
```

library(ggplot2)

```{r fig1, fig.path="figures/"}

pairs(learning2014[!names(learning2014) %in% c("gender")],col=learning2014$gender)

```
```{r datastructure}
summary(learning2014)
```

```{r fig2, fig.path="figures/", fig.dim=c(10,10), results='hide', message=FALSE}

library(GGally)

library(ggplot2)



# create a more advanced plot matrix with ggpairs()

ggpairs(learning2014, 

        mapping = aes(col = gender, alpha = 0.3), 

        lower = list(combo = wrap("facethist", bins = 20))

        )

```
```{r}

qplot(attitude, points, data = learning2014) + geom_smooth(method = "lm")

```
```{r, results='asis'}

my_model <- lm(points ~ attitude, data = learning2014)

results <- summary(my_model)
knitr::kable(results$coefficients, digits=3, caption="Regression coefficients")

```
```{r fig3, fig.path="figures/"}

plot(my_model, which=c(1,2,5))

```
